Symmetry-based constraints on the size of the dynamical Solar neighborhood

### 2019-04-09 paper/ms.tex
Some small edits throughout

Underlying the quantitative analysis of many of the mechanisms at work that
give rise to these signatures is the assumption of a global, axisymmetric
Galactocentric coordinate system \citep{2008gady.book.....B} in which the
Sun's position (and therefore the origin of the coordinate system) can be
defined and measured both precisely and accurately. In this system, the center
of the galaxy is usually taken to be the location of the central supermassive
black hole, Sgr~A\textsuperscript{*} \citep[e.g.][]{2004ApJ...616..872R}. From
stellar motions near Sgr~A\textsuperscript{*}, the distance from the Sun to
the Galactic center is taken to be $\sim 8.1\text{--}8.3\,\kpc$
\citep{2009ApJ...692.1075G,2018AA...615L..15G}. The Sun's vertical distance
from the Galactic midplane, the other independent measurement that sets this
coordinate system, can be determined by identifying where the stellar density
and velocities reach a maximum (effectively the median height of all disk
stars), and is usually taken to be $\sim 25\,\pc$ \citep{2001ApJ...553..184C},
with more recent measurements from {\em Gaia} DR2 placing it at $\sim 20\,\pc$
\citep{2019MNRAS.482.1417B}. Another strategy is to use the cold gas or HII
regions in the disk to define the Galactic midplane, leading to slightly
values ($\sim 5\,\pc$) for the Sun's relative height
\citep[e.g.][]{2019ApJ...871..145A}. A pre-{\em Gaia} review of these
measurements is given by \citet{2016ARAA..54..529B}.

Once a Galactocentric coordinate system has been established and a 6D phase
space measurement of a star has been made, it is often desirable to convert
this measurement into action space to concisely summarize its projected orbit,
model the stellar distribution function, or find stars with similar dynamical
properties. Actions are conserved quantities which describe the orbit of a
star under the assumption of regular, bound orbits in a system where the
equations of motion are separable in a particular coordinate system. They are
the typical integral of the Hamiltonian in cylindrical coordinates and are
given by,
\beq\label{eq:actions}
J_i \equiv
\frac{1}{2\pi} \oint_{\text{orbit}}p_i\,\text{d}x_i\text{,}
\eeq
where $i=R,z,\phi$ are the cylindrical radial, vertical, and azimuthal coordinates respectively and the $p_i$ are the conjugate momenta.\footnote{We typically write $J_r$ instead of $J_R$ for
aesthetic reasons.} In a slowly-evolving
axisymmetric potential, these actions are invariant and $J_{\phi} \equiv L_z$
\citep{2008gady.book.....B,2014RvMP...86....1S}.

With the advent of 6D phase-space measurements over a relatively large ($\gtrsim 2$ kpc) volume from the {\em Gaia} satellite, the study of stars' actions has gained new popularity. One reason is dimensionality reduction---an individual stellar orbit is concisely described by three actions, as opposed to six dimensions of phase space. Under the assumption of a phase-mixed system, the dynamical properties of a population of stars should be uniquely a function of their actions and independent of the conjugate angles, allowing investigation of the
relationship between {\em orbital} properties of stars and other intrinsic, and at least partially invariant, properties such as age or metallicity. Actions also provide a convenient basis for constructing models of the stellar distribution function, or for associating stars with similar dynamical properties (e.g. to determine membership in moving groups).

@@

In this work we discuss how the imposition of a global axisymmetric coordinate system based on a local measurement of the midplane location and Galactocentric distance introduces systematic error in the computation of actions, especially for stars in the thin disk. Our main point
is that the ``local'' Galactic midplane is not the same as the ``global''
Galactic midplane. For stars relatively far from the Sun,
the local midplane location will not agree with
our local midplane extrapolated onto their position, thanks to relatively small-scale variations in the density of the gas and stellar distribution. Converting positions and velocities of more distant stars from a heliocentric to a
Galactocentric coordinate system introduces a systematic bias in the $z$ coordinate, and hence the actions computed for these stars. The further from the solar neighborhood the target star, the more likely the mismatch will result in large systematic uncertainty, especially in the vertical action $J_z$. Depending on the type of orbit being studied, this sets a limit on the ``dynamical solar neighborhood'' over which the local coordinate frame can be usefully extended for action-space analysis. A similar argument applies to uncertainty in measurements of
$R_{\odot}$ and variations in the local standard of rest (LSR). In Section~\ref{sec:ref_frame}, we describe the general impact coordinate
system errors have on the measured actions.

Lacking an empirical, nonlocal definition of the global Galactic midplane or a measurement of the expected midplane variation in the Milky Way itself,  in Section~\ref{sec:local_fire} we attempt to quantify the azimuthal dependence of the
local midplane, and hence the expected size of the dynamical solar neighborhood, using cosmological-hydrodynamical zoom-in simulations of Milky Way-mass galaxies from the Feedback in Realistic Environments (FIRE)
collaboration\footnote{\url{https://fire.northwestern.edu}} \citep{2016ApJ...827L..23W,2017arXiv170206148H}.

@@

\section{Motivation} \label{sec:ref_frame}
We first demonstrate the significance to action computations, especially for
disk-like orbits, of systematic errors in the determination of the Galactic
midplane and distance to the Galactic center. The consequences we explore here
may also arise from various other systematic errors. For instance, the
axisymmetric Galactic potential model used in many works to compute actions
may not be a good description of the true potential --- or the parameters used
may yield a potential that is systematically incorrect outside an original
fitted region. In this section, we assume that the Galaxy is perfectly
described by our model axisymmetric potential, and simply explore the
consequences of systematic errors in the determination of the Galactocentric
coordinate system.

\subsection{The effect of midplane determination on orbit and action
calculations} \label{ssec:cartoon} We present a cartoon in
Figure~\ref{fig:cartoon} to show how an inaccurate determination of the
midplane leads to a phase-dependence of the actions calculated from an orbit
specified by a phase-space point and an assumed potential model. The $y$-axis
corresponds to the vertical height of the orbit and the $x$-axis orbital
phase. The solid gray line indicates the ``true'' orbit of the star; that is,
its orbit as it oscillates around the true midplane. The dashed gray line is
an erroneous midplane assigned to the potential model which an observer uses
to integrate the orbit of the star and hence estimate its actions. The model
potential is otherwise identical to the one in which the star is actually
moving.

@@

\subsection{Numerical methods} \label{ssec:action_comp}
We compute actions as in our previous work \citep{2018ApJ...867...31B}, using
the code \texttt{gala} v0.3 to perform orbit integrations and conversion to
action space \citep{2017JOSS....2..388P,Price-Whelan:2018}. To compute actions
we use the torus-mapping technique first presented by
\citet{1990MNRAS.244..634M} and adapted by \citet{2014MNRAS.441.3284S} to
calculate actions for an orbital time-series starting from a phase-space
position $(x, v)$ and integrated in a potential $\Phi$. We use the default
\texttt{MWPotential} as our potential, which is based on the Milky Way
potential available in \texttt{galpy} \citep{2015ApJS..216...29B}. This
potential includes a Hernquist bulge and nucleus \citep{1990ApJ...356..359H},
a Miyamoto-Nagai disk \citep{1975PASJ...27..533M}, and an NFW halo
\citep{1997ApJ...490..493N}, and is fit to empirically match some
observations. We use the Dormand-Prince 8(5,3) integration scheme
\citep{Dormand80:integrator} with a timestep of $1\,\Myr$ and integrate for
$5\,\Gyr$, corresponding to $\sim 20$ orbits for a Sun-like star.

Other methods for computing actions are used in the literature. For example,
the St\"ackel Fudge method \citep{2016MNRAS.457.2107S}, which uses a single
St\"ackel potential (with analytic actions) to approximate the Galactic
potential \citep{1985MNRAS.216..273D,2012MNRAS.426.1324B}, was used in many
recent works exploring actions in the Galactic disk
\citep[e.g.][]{2018arXiv180503653T,2018MNRAS.481.4093S,2018arXiv180803278T}.
For disk-like orbits, existing implementations of the St\"ackel Fudge method
are of acceptable accuracy, but since we also consider halo-like orbits in
this work we choose to use orbit integration and torus mapping throughout
\citep{2016MNRAS.457.2107S}.

\subsection{Quantification of the effect of systematic coordinate-system
errors} \label{ssec:quant}
First, we illustrate the phase-dependence induced by an
example systematic error in the Galactocentric coordinate system. We will
consider three orbits in the model potential described in
Section~\ref{ssec:action_comp} that are typical of stars in the stellar thin
disk, stellar thick disk, and the stellar halo. We summarize their initial
positions in phase space and the actions computed by integrating their orbits
in the correct potential in Table~\ref{tab:orbits}. Each orbit, integrated
without systematic coordinate errors, is plotted in Appendix~\ref{app:orbits}.
We will refer to these orbits by their names (thin-disk, thick-disk, halo)
henceforth.

We begin by considering the thick-disk orbit. Suppose that we could observe
the thick-disk star's phase-space position at many different times (and hence
different orbital phases), but used a coordinate system in which the midplane
is systematically offset in height by $100\,\pc$ from its actual location ---
i.e. we subtract the vector $(0, 0, 100)\,\pc$ from each position in the
orbit. This corresponds to an observer physically located at the position $(8,
0, 0)\,\kpc$ in the coordinate system of the true potential, but erroneously
thinking they are located at $(8, 0, 0.1)\,\kpc$.

@@

Starting from each observation of the star's phase-space position that such an
(extremely long-lived) observer makes every $\Myr$, we then compute the
actions as outlined above using an integration from that starting point in
phase space carried out using the true potential (but specifying the star's
present-day position using the systematically offset coordinate system). The
erroneous computed actions for each phase-space starting point are shown for
the first $\Gyr$ of the orbit in the {\em upper} panels
Figure~\ref{fig:one_orbit_wrong_ref}.\footnote{Occasionally the numerical
scheme fails and very large actions are reported by \texttt{gala}---we perform
a $5\sigma$ clip on each action to exclude such orbits, but this only excludes
a total of $5$ orbits out of the $1000$ considered for
Figure~\ref{fig:one_orbit_wrong_ref}. Some numerical artifacts remain, but the
vast majority of orbits are computed properly.}

@@

For a systematic offset in $z$ ({\em upper} panels), the
95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $2.2$ and
$6.2\,\actunit$ for $J_r$ and $J_z$, respectively. These are fractional errors
of $5.7\,\%$ and $85.7\,\%$, respectively. The error induced in $L_z$ is
negligible, as expected since $L_z$ is independent of $z$. It is worth
pointing out that a $100\,\pc$ error in an orbit with
$z_{\text{max}}=850\,\pc$ --- a $12\%$ error
--- induced a $43\%$ error in the computation of $J_z$.

For a systematic offset in $x$ (or Galactic center distance), the
95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $6.9$,
$47$, and $0.71 \actunit$ for $J_r$, $L_z$ and $J_z$, respectively. These are
fractional errors of $21\%$, $3.1\%$ and $3.1\%$, respectively.

@@

We now repeat the same procedure for systematic offsets between $0$ and
$500\,\pc$ in the $z$ and $x$ components. In
Figure~\ref{fig:many_orbit_wrong_ref}, we report the 95\uth minus 5\uth
percentile divided by the true action value for the three different fiducial
orbits in Table \ref{tab:orbits}. The thick-disk (\thickcolor) orbit is the
one from Figure~\ref{fig:one_orbit_wrong_ref}, but we also now consider the
effect on the action determined for the thin-disk (\thincolor) and halo
(\halocolor) orbits.

The top row of Figure~\ref{fig:many_orbit_wrong_ref} shows the spread induced
in each action for an offset in the $z$ component (i.e. the Galactic
midplane). In the {\em bottom row} we consider offsets in the $x$ component
(i.e. the Solar radius). The {\em left}, {\em middle}, and {\em right} columns
show the fractional spread in the values computed for $J_z$, $L_z$, and $J_r$,
respectively. We compute the fractional spread $\Delta J_i/J_i$ as the 95\uth
minus 5\uth percentile of the action computed using the offset coordinate
system over the course of the first $\Gyr$, divided by the true action
(computed in the correct coordinate system).

In the {\em upper middle} panel, there is essentially no spread in the
determination of $L_z$. This is expected since, in principle, $L_z$ only
depends on the $x$- and $y$-components of the position and velocity of the
stars\footnote{In practice, however, $L_z$ is computed as part of the
torus-fitting method.}, and is thus unaffected by offsets in $z$. Indeed, the
result we found in Figure~\ref{fig:one_orbit_wrong_ref} for the thick-disk
orbit holds for all orbit types.

The {\em upper right} panel shows that the fractional error in $J_z$ is more
exaggerated for more planar (disk-like) orbits. For the thin-disk orbit, a
systematic offset of $\sim30\,\pc$ in the $z$ coordinate already results in
$100\%$ deviations in the actions. The required offset for $100\%$ deviation
is $\sim225\,\pc$ for the thick-disk orbit. The halo orbit is relatively
resistant to errors in the midplane, with only $\sim25\%$ error in $J_z$ out
to a midplane offset of $500\,\pc$.

For the offset in the Solar radius {\em bottom panels}, the error is largest
for $J_r$, with some deviations resulting in $L_z$ and relatively small
deviations in $J_z$. In the {\em bottom middle} and {\em bottom right} panels
all three lines nearly overlap.

\begin{figure*}
\plotone{fig/schmactions_many_orbits.pdf}
\caption{We report the 95\uth minus 5\uth percentile of the error in the
measured action ($\Delta J_i$) from coordinate system errors for the thin,
thick, and halo orbits (Table~\ref{tab:orbits}). The {\em left}, {\em center},
and {\em right} panels show the result for $J_r$, $L_z$, and $J_z$,
respectively. The {\em upper} panels consider an offset in $z$ and the {\em
lower} panels consider an offset in $x$ (equivalently, an offset in the Solar
radius).}
\label{fig:many_orbit_wrong_ref}
\end{figure*}

@@

These simulations contain dark matter particles of mass $\sim35,000\,\Msun$,
gas particles of mass $\sim 7000$ to $20,000\,\Msun$, and star particles
of mass $\sim 5000 \text{--} 7000\,
\Msun$,\footnote{In fact, the simulation \mi{} contained a gas
splitting bug which causes a small number of gas and star particles to have
larger than typical masses. These constitute only $\sim0.2\%$ of the star
particles in the $\z=0$ snapshots, and are only a factor of a few more massive
than typical star particles. Since our results are broadly consistent between
the three simulations, we ignore this minor complication.} with the lower end
coming from stellar evolution \citep{2018arXiv180610564S}. Softening lengths
for dark matter and star particles are fixed at $112\,\pc$ and $11.2\,\pc$,
respectively.\footnote{This is $2.8$ times the often-quoted
Plummer-equivalent.} The gas softening length is adaptive, but at $z=0$ the
median softening length for cold ($T < 10^4\,\text{K}$) gas particles around
roughly solar positions (with cylindrical radii within $500\,\pc$ of
$8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is $12.4, 12.5, 10.3\,\pc$ for \mi{},
\mf{}, and \mm{}, respectively. These values are summarized in
Table~\ref{tab:scale_height}.

>>>>

Underlying the quantitative analysis of many of the mechanisms at work that
give rise to these signatures is the assumption of a global, axisymmetric
Galactocentric coordinate system \citep{2008gady.book.....B} in which the
Sun's position (and therefore the origin of the coordinate system) can be
defined and measured both precisely and accurately. In this system, the center
of the Galaxy is usually taken to be the location of the central supermassive
black hole, Sgr~A\textsuperscript{*} \citep[e.g.][]{2004ApJ...616..872R}. From
stellar motions near Sgr~A\textsuperscript{*}, the distance from the Sun to
the Galactic center is taken to be $\sim 8.1\text{--}8.3\,\kpc$
\citep{2009ApJ...692.1075G,2018AA...615L..15G}. The Sun's vertical distance
from the Galactic midplane, the other independent measurement that sets this
coordinate system, can be determined by identifying where the stellar density
and velocities reach a maximum (effectively the median height of all disk
stars), and is usually taken to be $\sim 25\,\pc$ \citep{2001ApJ...553..184C},
with more recent measurements from {\em Gaia} DR2 placing it at $\sim 20\,\pc$
\citep{2019MNRAS.482.1417B}. Another strategy is to use the cold gas or HII
regions in the disk to define the Galactic midplane, leading to slightly
different values ($\sim 5\,\pc$) for the Sun's relative height
\citep[e.g.][]{2019ApJ...871..145A}. A pre-{\em Gaia} review of these
measurements is given by \citet{2016ARAA..54..529B}.

Once a Galactocentric coordinate system has been established and a 6D phase
space measurement of a star has been made, it is often desirable to convert
this measurement into action space to concisely summarize its projected orbit,
model the stellar distribution function, or find stars with similar dynamical
properties. Actions are conserved quantities which describe the orbit of a
star under the assumption of regular, bound orbits in a system where the
equations of motion are separable in a particular coordinate system. They are
the typical integral of the Hamiltonian in cylindrical coordinates and are
given by,
\beq\label{eq:actions}
J_i \equiv
\frac{1}{2\pi} \oint_{\text{orbit}}p_i\,\text{d}x_i\text{,}
\eeq
where $i=R,z,\phi$ are the radial, vertical, and azimuthal coordinates respectively in a cylindrical coordinate system and the $p_i$ are the conjugate momenta.\footnote{We typically write $J_r$ instead of $J_R$ for
aesthetic reasons.} In a slowly-evolving
axisymmetric potential, these actions are invariant and $J_{\phi} \equiv L_z$
\citep{2008gady.book.....B,2014RvMP...86....1S}.

With the advent of 6D phase-space measurements over a relatively large ($\gtrsim 2$ kpc) volume from the {\em Gaia} satellite, the study of stars' actions has gained new popularity. One reason is dimensionality reduction---an individual stellar orbit is concisely described by three actions, as opposed to six dimensions of phase space. Second, under the assumption of a phase-mixed system, the dynamical properties of a population of stars should be uniquely a function of their actions and independent of the conjugate angles, allowing investigation of the
relationship between {\em orbital} properties of stars and other intrinsic, and at least partially invariant, properties such as age or metallicity. Actions also provide a convenient basis for constructing models of the stellar distribution function, or for associating stars with similar dynamical properties (e.g. to determine membership in moving groups).

@@

In this work we discuss how the imposition of a global axisymmetric coordinate system based on a local measurement of the midplane location and Galactocentric distance introduces systematic error in the computation of actions, especially for stars in the thin disk. Our main point
is that the ``local'' Galactic midplane is not the same as the ``global''
Galactic midplane. For stars relatively far from the Sun,
the local midplane location will not agree with
our local midplane extrapolated onto their position, thanks to relatively small-scale variations in the density of the gas and stellar distribution. Converting positions and velocities of more distant stars from a heliocentric to a
Galactocentric coordinate system introduces a systematic bias in the $z$ coordinate, and hence the actions computed for these stars. The further from the solar neighborhood the target star, the more likely the mismatch will result in large systematic uncertainty, especially in the vertical action $J_z$. Depending on the type of orbit being studied, this sets a limit on the ``dynamical solar neighborhood'' over which the local coordinate frame can be usefully extended for action-space analysis. A similar argument applies to uncertainty in measurements of
$R_{\odot}$\footnote{Here, and throughout, we use $R_{\odot}$ to mean the distance from the Sun to the Galactic center.} and variations in the local standard of rest (LSR).

In Section~\ref{sec:ref_frame}, we describe the general impact coordinate
system errors have on the measured actions.
Lacking an empirical, nonlocal definition of the global Galactic midplane or a measurement of the expected midplane variation in the Milky Way itself,  in Section~\ref{sec:local_fire} we attempt to quantify the azimuthal dependence of the
local midplane, and hence the expected size of the dynamical solar neighborhood, using cosmological-hydrodynamical zoom-in simulations of Milky Way-mass galaxies from the Feedback in Realistic Environments (FIRE)
collaboration\footnote{\url{https://fire.northwestern.edu}} \citep{2016ApJ...827L..23W,2017arXiv170206148H}.

@@

\section{Motivation} \label{sec:ref_frame}
We first demonstrate the significance of systematic errors in the determination of the Galactic midplane and distance to the Galactic center to action computations. We will find that such errors are especially important for disk-like orbits.
The consequences we explore here
may also arise from various other systematic errors. For instance, the
axisymmetric Galactic potential model used in many works to compute actions
may not be a good description of the true potential --- or the parameters used
may yield a potential that is systematically incorrect outside an original
fitted region. In this work, we assume that the Galaxy is perfectly
described by our model axisymmetric potential, and simply explore the
consequences of systematic errors in the determination of the Galactocentric
coordinate system.

\subsection{Cartoon effect of midplane error} \label{ssec:cartoon}
We present a cartoon in
Figure~\ref{fig:cartoon} to show how an inaccurate determination of the
midplane leads to a phase-dependence of the actions calculated from an orbit
specified by a phase-space point and an assumed potential model. The $y$-axis
corresponds to the vertical height of the orbit and the $x$-axis orbital
phase. The solid gray line indicates the ``true'' orbit of the star; that is,
its orbit as it oscillates around the true midplane. The dashed gray line is
an erroneous midplane assigned to the potential model which an observer uses
to integrate the orbit of the star and hence estimate its actions. The model
potential is otherwise identical to the one in which the star is actually
moving.

@@

\subsection{Numerical methods} \label{ssec:action_comp}
We now move on to quantifying the argument made in Section~\ref{ssec:cartoon}.
We compute actions as in our previous work \citep{2018ApJ...867...31B}, using
the code \texttt{gala} v0.3 to perform orbit integrations and conversion to
action space \citep{2017JOSS....2..388P,Price-Whelan:2018}. To compute actions
we use the torus-mapping technique first presented by
\citet{1990MNRAS.244..634M} and adapted by \citet{2014MNRAS.441.3284S} to
calculate actions for an orbital time-series starting from a phase-space
position $(x, v)$ and integrated in a potential $\Phi$. We use the default
\texttt{MWPotential} as our potential, which is based on the Milky Way
potential available in \texttt{galpy} \citep{2015ApJS..216...29B}. This
potential includes a Hernquist bulge and nucleus \citep{1990ApJ...356..359H},
a Miyamoto-Nagai disk \citep{1975PASJ...27..533M}, and an NFW halo
\citep{1997ApJ...490..493N}, and is fit to empirically match some
observations. We use the Dormand-Prince 8(5,3) integration scheme
\citep{Dormand80:integrator} with a timestep of $1\,\Myr$ and integrate for
$5\,\Gyr$, corresponding to $\sim 20$ orbits for a Sun-like star.

Other methods for computing actions are used in the literature. For example,
the St\"ackel Fudge method \citep{2016MNRAS.457.2107S}, which uses a single
St\"ackel potential (with analytic actions) to approximate the Galactic
potential \citep{1985MNRAS.216..273D,2012MNRAS.426.1324B}, was used in many
recent works exploring actions in the Galactic disk
\citep[e.g.][]{2018arXiv180503653T,2018MNRAS.481.4093S,2018arXiv180803278T}.
For disk-like orbits, existing implementations of the St\"ackel Fudge method
are of acceptable accuracy, but since we also consider halo-like orbits in
this work (where the St\"ackel Fudge method is inaccurate) we choose to use orbit integration and torus mapping throughout
\citep{2016MNRAS.457.2107S}.

\subsection{Quantification of the midplane effect} \label{ssec:quant}
We quantify how a systematic error in the Galactocentric coordinate system induces phase-dependence in the {\em observed} actions.
We will
consider three orbits in the model potential described in
Section~\ref{ssec:action_comp} that are typical of stars in the stellar thin
disk, stellar thick disk, and the stellar halo. We summarize their initial
positions in phase space and the actions computed by integrating their orbits
in the correct potential in Table~\ref{tab:orbits}. Each orbit, integrated
without systematic coordinate errors, is plotted in Appendix~\ref{app:orbits}.
We will refer to these orbits by their names (thin-disk, thick-disk, halo)
henceforth.

We begin by considering the thick-disk orbit. Suppose that we could observe
the thick-disk star's phase-space position at many different times (and hence
different orbital phases), but used a coordinate system in which the midplane
is systematically offset in height by $100\,\pc$ from its actual location ---
i.e. we subtract the vector $(0, 0, 100)\,\pc$ from each position in the
orbit. This corresponds to an observer physically located at e.g. the position $(8,
0, 0)\,\kpc$ in the coordinate system of the true potential, but erroneously
thinking they are located at $(8, 0, 0.1)\,\kpc$.

@@

Starting from each observation of the star's phase-space position that such an
(extremely long-lived) observer makes every $\Myr$, we then compute the
actions as outlined above using an integration from that starting point in
phase space carried out using the true potential but specifying the star's
present-day position using the systematically offset coordinate system. Essentially we are shifting the integrated orbit and then reintegrating at each point along the shifted orbit, where the shifted orbit is the observed orbit. The
erroneous computed actions for each phase-space starting point are shown for
the first $\Gyr$ of the orbit in the {\em upper} panels
Figure~\ref{fig:one_orbit_wrong_ref}.\footnote{Occasionally the numerical
scheme fails and very large actions are reported by \texttt{gala}---we perform
a $5\sigma$ clip on each action to exclude such orbits, but this only excludes
a total of $5$ orbits out of the $1000$ considered for
Figure~\ref{fig:one_orbit_wrong_ref}. Some numerical artifacts remain, but the
vast majority of orbits are computed properly.}

@@

For a systematic offset in $z$ ({\em upper} panels), the
95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $2.2$ and
$6.2\,\actunit$ for $J_r$ and $J_z$, respectively. These are fractional errors
of $5.7\,\%$ and $85.7\,\%$, respectively. The error induced in $L_z$ is
negligible, as expected since $L_z$ is independent of $z$. It is worth
pointing out that a $100\,\pc$ error in an orbit with
$z_{\text{max}}=850\,\pc$ --- a $12\%$ error
--- induced an $86\%$ error in the computation of $J_z$.

For a systematic offset in $x$ (or Galactic center distance), the
95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $6.9$,
$47$, and $0.71\,\actunit$ for $J_r$, $L_z$ and $J_z$, respectively. These are
fractional errors of $21\%$, $3.1\%$ and $3.1\%$, respectively.

@@

We now repeat the same procedure for systematic offsets between $0$ and
$500\,\pc$ in the $z$ and $x$ components. In
Figure~\ref{fig:many_orbit_wrong_ref}, we report the 95\uth{} minus 5\uth{}
percentile divided by the true action value for the three different fiducial
orbits in Table \ref{tab:orbits}. The thick-disk (\thickcolor) orbit is the
one from Figure~\ref{fig:one_orbit_wrong_ref}, but we also now consider the
effect on the action determined for the thin-disk (\thincolor) and halo
(\halocolor) orbits.

The top row of Figure~\ref{fig:many_orbit_wrong_ref} shows the spread induced
in each action for an offset in the $z$-component (i.e. the Galactic
midplane). In the {\em bottom row} we consider offsets in the $x$ component
(i.e. the Solar radius). The {\em left}, {\em middle}, and {\em right} columns
show the fractional spread in the values computed for $J_z$, $L_z$, and $J_r$,
respectively. We compute the fractional spread $\Delta J_i/J_i$ as the 95\uth{}
minus 5\uth{} percentile of the action computed using the offset coordinate
system over the course of the first $\Gyr$, divided by the true action
(computed in the correct coordinate system).

In the {\em upper middle} panel of Figure~\ref{fig:many_orbit_wrong_ref}, there is essentially no spread in the
determination of $L_z$. This is expected since, in principle, $L_z$ only
depends on the $x$- and $y$-components of the position and velocity of the
stars\footnote{In practice, however, $L_z$ is computed as part of the
torus-fitting method.}, and is thus unaffected by offsets in $z$. Indeed, the
result we found in Figure~\ref{fig:one_orbit_wrong_ref} for the thick-disk
orbit holds for all orbit types.

The {\em upper right} panel of Figure~\ref{fig:many_orbit_wrong_ref} shows that the fractional error in $J_z$ is more
exaggerated for more planar (disk-like) orbits. For the thin-disk orbit, a
systematic offset of $\sim30\,\pc$ in the $z$ coordinate already results in
$100\%$ deviations in the actions. The required offset for $100\%$ deviation
is $\sim225\,\pc$ for the thick-disk orbit. The halo orbit is relatively
resistant to errors in the midplane, with only $\sim25\%$ error in $J_z$ out
to a midplane offset of $500\,\pc$.

For the offset in the Solar radius ({\em bottom panels}), the error is largest
for $J_r$, with some deviations resulting in $L_z$ and relatively small
deviations in $J_z$. In the {\em bottom middle} and {\em bottom right} panels
all three lines nearly overlap.

\begin{figure*}
\plotone{fig/schmactions_many_orbits.pdf}
\caption{We report the 95\uth{} minus 5\uth{} percentile of the error in the
measured action ($\Delta J_i$) from coordinate system errors for the thin,
thick, and halo orbits (Table~\ref{tab:orbits}). The {\em left}, {\em center},
and {\em right} panels show the result for $J_r$, $L_z$, and $J_z$,
respectively. The {\em upper} panels consider an offset in $z$ and the {\em
lower} panels consider an offset in $x$ (equivalently, an offset in the Solar
radius).}
\label{fig:many_orbit_wrong_ref}
\end{figure*}

@@

These simulations contain dark matter particles of mass $\sim35,000\,\Msun$,
gas particles of mass $\sim 7000$ to $20,000\,\Msun$, and star particles
of mass $\sim 5000 \text{--} 7000\,
\Msun$,\footnote{In fact, the simulation \mi{} contained a gas
splitting bug which causes a small number of gas and star particles to have
larger than typical masses. These constitute only $\sim0.2\%$ of the star
particles in the $\z=0$ snapshots, and are only a factor of a few more massive
than typical star particles. Since our results are broadly consistent between
the three simulations, we ignore this minor complication.} with the lower end
coming from stellar evolution \citep{2018arXiv180610564S}. Softening lengths
for dark matter and star particles are fixed at $112\,\pc$ and $11.2\,\pc$,
respectively.\footnote{This is $2.8$ times the often-quoted
Plummer-equivalent.} The gas softening length is adaptive, but at $\z=0$ the
median softening length for cold ($T < 10^4\,\text{K}$) gas particles around
roughly solar positions (with cylindrical radii within $500\,\pc$ of
$8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is $12.4, 12.5, 10.3\,\pc$ for \mi{},
\mf{}, and \mm{}, respectively. These values are summarized in
Table~\ref{tab:scale_height}.
